Reasoning About a Machine with Local Capabilities

### 2019-02-04 toplas/paper.tex
Update the introduction

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) and/or rely on techniques like
machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
verification~\citep{morrisett_system_1999}, virtual machines with a
native stack~\citep{lindholm_java_2014} or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

@@

Unlike commodity processors, capability machines lend themselves well to
enforcing local-state encapsulation. Potentially, they will enable compilation
schemes that enforce this property in an efficient but also 100\% watertight way
(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
in a new attack-defense arms race). However, a lot needs to happen before we get
there. For example, it is far from trivial to devise a compilation scheme
adapted to the details of a specific source language's notion of encapsulation
(e.g., private member variables in OO languages often behave quite differently
than private state in ML-like languages). And even if such a scheme were
defined, a formal proof depends on a formalization of the encapsulation provided
by the capability machine at hand.

@@

\begin{toplas}
  This paper is an extension of a published conference paper~\citep{Skorstengaard:esop18}.
  We have made minor improvements to readability and completeness throughout the paper, but we highlight the following changes:
  \begin{itemize}
  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
  \item We have added proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
  \item We have added figures that illustrate central parts of the calling convention.
  \item We have added a section on malloc.
    Specifically, we provide the specification for the malloc used in the examples of the paper.
  \item We have added a section on macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
  \item We have added a section on reasoning about programs that run on a capability machine.
    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
    It also introduces a number of lemmas that prove recurring bits once and for all.
  \item We have added many details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
    Further, many of the details we did still not include in the paper are now in the appendix of this paper.
  \end{itemize}
\end{toplas}

>>>>

Compromising software security is often based on attacks that break programming language properties relied upon by software authors, such as control-flow correctness, local-state encapsulation, etc.
Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-mitigation arms race) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
The latter techniques essentially emulate protection techniques on existing hardware, at the cost of performance, system complexity and/or security.

@@

Unlike commodity processors, capability machines lend themselves well to
enforcing local-state encapsulation. Potentially, they will enable compilation
schemes that enforce this property in an efficient but also 100\% watertight way
(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
in a new attack-mitigation arms race). However, a lot needs to happen before we get
there. For example, it is far from trivial to devise a compilation scheme
adapted to the details of a specific source language's notion of encapsulation
(e.g., private member variables in OO languages often behave quite differently
than private state in ML-like languages). And even if such a scheme were
defined, a formal proof depends on a formalization of the encapsulation provided
by the capability machine at hand.

@@

\begin{toplas}
  This paper is an extension of the published conference paper~\citet{Skorstengaard:esop18}.
  We have made improvements to readability and completeness throughout the paper, but we highlight the following changes:
  \begin{itemize}
  \item An introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
  \item Proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
  \item Figures that illustrate central parts of the calling convention.
  \item A section about malloc.
    Specifically, we provide the specification for the malloc used in the examples of the paper.
  \item A section about macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
  \item A section on reasoning about programs that run on a capability machine.
    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
    It also introduces a number of lemmas that prove recurring bits once and for all.
  \item Details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
    Further, many of the still omitted details have been added to the paper's appendix.
  \end{itemize}
\end{toplas}
